Methods that remove batch effects while retaining group differences may lead to exaggerated confidence in downstream analyses

### 2014-07-24 tex/manuscript.tex
forandret litt på figure1

\begin{figure}[!p]
\centering\includegraphics[width=13cm]{Fig/boxplots.pdf}
%\centering\includegraphics[width=13cm]{Fig/boxplots.png}
\caption{We simulated expression of one gene from four study groups unevenly distributed in 3 batches containing 50+0+0+20, 0+50+0+20 and 0+0+50+20 samples from groups 1 to 4. This case, design and effects, was selected to illustrate the spurious effects that may arise from different batch adjustments. The Y-axis represent the expression, while the X-axis is used to better visually differentiate between the batches. Boxes showing group mean with 95\% confidence intervals are displayed to the right of the measurements.
a) The ``true'' values where groups 1 and 2 have lower means than groups 3 and 4.
b) After adding batch effects, all the groups seem different.
c) After batch centering, group 3 and 4 seem to differ while group 1 and 3 are more similar.
d) After two-way ANOVA based batch centering (using \texttt{limma}), group 1 and 2 seem to differ.
e) The least squares estimates of the group means from a two-way ANOVA which represent a more reliable way of estimating group differences than the two-step procedure in c) and d).
}
\label{fig:boxplots}
\end{figure}

>>>>

\begin{figure}[!p]
\centering\includegraphics[width=13cm]{Fig/boxplots.pdf}
%\centering\includegraphics[width=13cm]{Fig/boxplots.png}
\caption{We simulated expression of one gene from four study groups unevenly distributed in 3 batches containing 50+0+0+20, 0+50+0+20 and 0+0+50+20 samples from groups 1 to 4. This case, design and effects, was selected to illustrate the spurious effects that may arise from different batch adjustments. The Y-axis represent the expression, while the X-axis is used to better visually differentiate between the batches. Circles, triangles and crosses indicate the measurment of each sample and which batch it is part of. Boxes showing group mean with 95\% confidence intervals are displayed to the right of the measurements and are ment to imply if a statistical test would find significant group differences, true or false.
a) The ``true'' values measured in a system without batch effects. Groups 1 and 2 have lower means than groups 3 and 4.
b) The same samples as in a) but measured in a system with batch effects. All the groups seem different.
c) The measurements in b) are adjusted with batch centering. Group 3 and 4 seem to differ while group 1 and 3 are more similar.
d) The measurements in b) are adjusted with two-way ANOVA based batch centering (using \texttt{limma}). Group 1 and 2 seem to differ.
e) The least squares estimates of the group means from a two-way ANOVA which represent a more reliable way of estimating group differences of the measurements in b) than the two-step procedure in c) and d).
}
\label{fig:boxplots}
\end{figure}
